On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-08-19 Manuscript/manuscript.tex
Manuscript Update

Block bootstrap is a tool to construct confidence intervals (CI)
to make inferences about dependent data. Early ideas were developed
independently by 
\citet{hall1985resampling, carlstein1986use,ckunsch1989jackknife}.
It has since been applied in various fields such as econometrics
\citep{mackinnon2006bootstrap} and meteorology \citep{varga2017generalised}.
Block bootstrap is useful for serially dependent
data when the serial dependence is not specified or not of primary interest.
The method is expected to produce CIs with coverage rates
matching their nominal levels as the sample size grows. However,
when dealing with finite sample sizes, an important question is how large the
sample size must be for block bootstrap CIs to have the
desired coverage rates.

@@

Suppose that we have repeated the steps in the last subsection $B$ times, and
that for $b \in \{1, \ldots, B\}$, we have obtained an estimator
$\hat\theta_n^{(b)}$ based on the $b$th bootstrap sample using the same
method
that was applied to $\{X_t: t = 1, \ldots, n\}$ to obtain $\hat\theta_n$.
Now the question is, how to construct a CI for $\theta$
using the $B$ bootstrap point estimates
$\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$.
We consider five kinds of CIs.

\paragraph{Standard Normal CI}
Assuming that $\hat\theta_n$ is asymptotically normally distributed with
$\theta$ as the mean, we just need an estimate of the standard error to
construct an approximate CI \citep[p.168]{efron1993introduction}.
Let $\widehat{\text{SE}}$ be the empirical standard error of the block bootstrap
copies of $\hat\theta_n^{(b)}$, $b = 1, \ldots, B$. Let $z_{(\alpha)}$ be the
upper $\alpha$ quantile of the standard normal distribution.
A $(1 - \alpha)100\%$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - z_{(\alpha/2)}\widehat{\text{SE}}).
\]
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Its validity relies on whether the distribution of $\hat\theta_n$ is reasonably
well approximated by its asymptotic normal distribution and whether the
bootstrap $\widehat{\text{SE}}$ approximates the true standard error.

@@

\paragraph{Bias Corrected and Accelerated ($BC_a$) Bootstrap CI}
The $BC_a$ CI was first suggested in \citet{efron1987better}.
Let $\hat{z}_0 = \Phi^{-1}\{\#\{\hat\theta_n^{(b)} < \hat{\theta}_n\} / B\}$
for $b \in \{1, \ldots, B\}$. 
Let $Z_{(i)}$ be the original sample without the $i$th block $z_i$, let
$\hat{\theta}_{(i)}$ be the jackknife values of $Z_{(i)}$
for $i \in \{1, \ldots, n\}$, 
and let $\hat{\theta}_{(.)} = \sum_{i=1}^{k} \hat{\theta}_{(i)} / k$. 
Let 
\[
\hat{a} = \frac{\sum_{i=1}^{k} (\hat{\theta}_{(.)} -
  \hat{\theta}_{(i)})^3}{6\{\sum_{i=1}^{k} (\hat{\theta}_{(.)} -
  \hat{\theta}_{(i)})^2\}^{3/2}}.
\]
Define
\[
\alpha_1 = \Phi\left(z_0 + \frac{z_{0} +
  z_{\alpha/2}}{1 - \hat{a}(z_{0} + z_{\alpha/2})}\right)
\text{ and }
\alpha_2 = \Phi\left(z_0 + \frac{z_{0} +
  z_{1 - \alpha/2}}{1 - \hat{a}(z_{0} + z_{1 - \alpha/2})}\right).
\]
A $(1 - \alpha)100\%~BC_a$ CI is
\[
(\hat\theta_{n, \alpha_1}^{(b)}, \quad \hat\theta_{n, \alpha_2}^{(b)}).
\]
This CI is not necessarily centered by the point estimator $\hat\theta_n$. The
$BC_a$ method corrects for bias and skewness of the $B$ bootstrap point
estimates $\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$ by including
bias-correction and acceleration factors. The acceleration factor refers to
the 
rate of change of the standard error of $\hat\theta_n$ with respect to
$\theta$.

\paragraph{A Proposed CI}
We propose a CI that is centered at the point estimator and uses the variation
in the bootstrap estimates to construct the error bound. This interval requires
the computation of $\bar\theta_n^{(b)}$, the mean of all bootstrap point
estimates, and $\widehat{\text{bias}} = \bar\theta_n^{(b)} - \hat\theta_n$.
A $(1 - \alpha)100\%$ CI is
\[
(\hat\theta_{n, \alpha/2}^{(b)} - \widehat{\text{bias}}, \quad
\hat\theta_{n, 1 - \alpha/2}^{(b)} - \widehat{\text{bias}}).
\]
This CI is centered around $\hat\theta_n$ and can also be rewritten
as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{(b)} - \bar\theta_n^{(b)}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{(b)} - \bar\theta_n^{(b)}).
\]
It is asymmetric, as different critical values are used to compute
the lower and upper bounds.
\jy{Double check}

>>>>

Block bootstrap is a tool to construct confidence intervals (CI)
to make inferences about dependent data. Early ideas were developed
independently by 
\citet{hall1985resampling, carlstein1986use,kunsch1989jackknife}.
It has since been applied in various fields such as econometrics
\citep{mackinnon2006bootstrap} and meteorology \citep{varga2017generalised}.
Block bootstrap is useful for serially dependent
data when the serial dependence is not specified or not of primary interest.
The method is expected to produce CIs with coverage rates
matching their nominal levels as the sample size grows. However,
when dealing with finite sample sizes, an important question is how large the
sample size must be for block bootstrap CIs to have the
desired coverage rates.

@@

Suppose that we have repeated the steps in the last subsection $B$ times, and
that for $b \in \{1, \ldots, B\}$, we have obtained an estimator
$\hat\theta_n^{(b)}$ based on the $b$th bootstrap sample using the same
method
that was applied to $\{X_t: t = 1, \ldots, n\}$ to obtain $\hat\theta_n$.
Now the question is how to construct a CI for $\theta$
using the $B$ bootstrap point estimates
$\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$.
We consider six kinds of CIs.

\paragraph{Standard Normal CI}
Assuming that $\hat\theta_n$ is asymptotically normally distributed with
$\theta$ as the mean, we just need an estimate of the standard error to
construct an approximate CI \citep[p.168]{efron1993introduction}.
Let $\widehat{\text{SE}}$ be the empirical standard error of the block bootstrap
copies of $\hat\theta_n^{(b)}$, $b = 1, \ldots, B$. Let $z_{(\alpha)}$ be the
upper $\alpha$ quantile of the standard normal distribution.
A $(1 - \alpha)100\%$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - z_{(\alpha/2)}\widehat{\text{SE}}).
\]
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric. The
Standard CI is a table-based bootstrap interval
approach. Its validity relies on whether the distribution of $\hat\theta_n$ is
 reasonably
well approximated by its asymptotic normal distribution and whether the
bootstrap $\widehat{\text{SE}}$ approximates the true standard error.

@@

\paragraph{Bias Corrected and Accelerated ($BC_a$) Bootstrap CI}
The $BC_a$ CI was first suggested in \citet{efron1987better}.
Let $\hat{z}_0 = \Phi^{-1}\{\#\{\hat\theta_n^{(b)} < \hat{\theta}_n\} / B\}$
for $b \in \{1, \ldots, B\}$. 
Let $Z_{(i)}$ be the original sample without the $i$th block $z_i$ for 
$i \in \{1, \ldots, k\}$, let
$\hat{\theta}_{(i)}$ be the statistic of $Z_{(i)}$,
and let $\hat{\theta}_{(.)} = \sum_{i=1}^{k} \hat{\theta}_{(i)} / k$. 
Let 
\[
\hat{a} = \frac{\sum_{i=1}^{k} (\hat{\theta}_{(.)} -
  \hat{\theta}_{(i)})^3}{6\{\sum_{i=1}^{k} (\hat{\theta}_{(.)} -
  \hat{\theta}_{(i)})^2\}^{3/2}}.
\]
Define
\[
\alpha_1 = \Phi\left(z_0 + \frac{z_{0} +
  z_{\alpha/2}}{1 - \hat{a}(\hat{z}_{0} + z_{\alpha/2})}\right)
\text{ and }
\alpha_2 = \Phi\left(z_0 + \frac{z_{0} +
  z_{1 - \alpha/2}}{1 - \hat{a}(\hat{z}_{0} + z_{1 - \alpha/2})}\right).
\]
A $(1 - \alpha)100\%~BC_a$ CI is
\[
(\hat\theta_{n, \alpha_1}^{(b)}, \quad \hat\theta_{n, \alpha_2}^{(b)}).
\]
This CI is not necessarily centered by the point estimator $\hat\theta_n$. The
$BC_a$ method corrects for bias and skewness of the $B$ bootstrap point
estimates $\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$ by including
bias-correction and acceleration factors. The acceleration factor refers to
the 
rate of change of the standard error of $\hat\theta_n$ with respect to
$\theta$.

\paragraph{A Proposed CI}
We propose a CI that is centered at the point estimator and uses the variation
in the bootstrap estimates to construct the error bound. This interval requires
the computation of $\bar\theta_n^{(b)}$, the mean of all bootstrap point
estimates, and $\widehat{\text{bias}} = \bar\theta_n^{(b)} - \hat\theta_n$.
A $(1 - \alpha)100\%$ CI is
\[
(\hat\theta_{n, \alpha/2}^{(b)} - \widehat{\text{bias}}, \quad
\hat\theta_{n, 1 - \alpha/2}^{(b)} - \widehat{\text{bias}}).
\]
This CI is centered around $\hat\theta_n$ and can also be rewritten
as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{(b)} - \bar\theta_n^{(b)}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{(b)} - \bar\theta_n^{(b)}).
\]
It is asymmetric, as different critical values are used to compute
the lower and upper bounds.
\jy{Double check} \mc{changed BCA, had to rerun simulations because I did k
degrees of freedom instead of k - 1}
